---
title: "First Step Visualization"
author: "Ignacio Almodóvar Cárdenas, Javier Muñoz Flores & Luis Ángel Rodríguez García"
date: "22-04-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Load the libraries:

```{r, results='hide'}
library(igraph)
```

Load the database within the `bicimad-data.RData`file:

```{r}
load("bicimad-data.RData")
```

First all, we are going to create 5 different zones of the vertices based on the latitude and the longitude of each vertex. Every zone represents a geographical area of Madrid: northwest (`NW`), northeast (`NE`), center (`C`), southwest (`SW`) and southeast (`SE`).

```{r}
library(dplyr)
library(RColorBrewer)

pal <- brewer.pal(n = 5, name = "Dark2")

coordinates <- stations %>% select(c(4,5)) %>% 
  mutate(
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude))
longitude_quantiles <- quantile(as.numeric(stations$longitude), prob=c(.30,.70))
latitude_quantiles <- quantile(as.numeric(stations$latitude), prob=c(.30,.70))
areas <- ifelse(
  Reduce("&", list(
    as.numeric(stations$latitude) > latitude_quantiles[1], 
    as.numeric(stations$latitude) < latitude_quantiles[2], 
    as.numeric(stations$longitude) > longitude_quantiles[1], 
    as.numeric(stations$longitude) < longitude_quantiles[2])), "C", 
                ifelse(as.numeric(stations$longitude) >= longitude_quantiles[2], 
                ifelse(as.numeric(stations$latitude) >= latitude_quantiles[2], 'NE', 'SE'),
                ifelse(as.numeric(stations$latitude) >= latitude_quantiles[2], 'NW', 'SW')))
# NE: "#1B9E77"
# SE: "#D95F02"
# SW: "#7570B3"
# NW: "#E7298A"
# C:  "#66A61E"
V(bicimad)$area <- areas
V(bicimad)$color <- sapply(areas, function(x) 
    ifelse(x=="NE", pal[1],
           ifelse(x=="SE", pal[2], 
                  ifelse(x=="SW", pal[3], 
                         ifelse(x=="NW", pal[4], pal[5]))))
    )
E(bicimad)$color <- tail_of(bicimad, E(bicimad))$color
bicimad <- delete_vertex_attr(bicimad, 'color')
```

# Visualization of the network

## Simple graph layouts

### Circular layout

```{r, fig.width=12, fig.height=12}
plot.igraph(bicimad, layout=layout_in_circle, main="Bicimad network with circular layout", vertex.size=4, edge.width=0.2, edge.arrow.size=0.3, edge.arrow.width=1, vertex.label.cex=0.5)
```

### Circular layout

```{r, fig.width=12, fig.height=12}
plot.igraph(bicimad, layout=layout_as_star,main="Bicimad network with star layout", vertex.size=4, edge.width=0.2, edge.arrow.size=0.3, edge.arrow.width=1, vertex.label.cex=0.5)
```

### Tree layout

```{r, fig.width=12, fig.height=12}
plot.igraph(bicimad, layout=layout_as_tree, main="Bicimad network with tree layout", vertex.size=4, edge.width=0.2, edge.arrow.size=0.3, edge.arrow.width=1, vertex.label.cex=0.5)
```

### Grid layout

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_on_grid, main="Bicimad network with grid layout", vertex.size=4, edge.width=0.2, edge.arrow.size=0.3, edge.arrow.width=1, vertex.label.cex=0.5)
```

### Sphere layout

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_on_sphere, main="Bicimad network with sphere layout", vertex.size=4, edge.width=0.2, edge.arrow.size=0.3, edge.arrow.width=1, vertex.label.cex=0.5)
```

### Chord diagram

In this case, Chord diagram is not useful as the graph contains multiple edges.

## Graph layouts based on maximizing energy functions

### Davidson-Harel layout

This layout takes too while to be executed, this might happen because the graph contains multiple edges

```{r}
plot.igraph(bicimad, layout=layout_with_dh, main="Bicimad network with Davidson-Harel layout", vertex.size=8, vertex.label.cex=0.6)
```


### Fruchterman-Reingold layout

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_with_fr, main="Bicimad network with Fruchterman-Reingold layout", vertex.size=8, vertex.label.cex=0.6)
```

### GEM force-directed layout

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_with_gem, main="Bicimad network with GEM force-directed layout", vertex.size=8, vertex.label.cex=0.6)
```

### Graphopt layout

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_with_graphopt, main="Bicimad network with graphopt layout", vertex.size=8, vertex.label.cex=0.6)
```

### Kamada-Kawai layout

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_with_kk, main="Bicimad network with Kamada-Kawai layout", vertex.size=8, vertex.label.cex=0.6)
```

## Graph layouts based on multidimensional scaling

```{r, fig.width=15, fig.height=15}
plot.igraph(bicimad, layout=layout_with_mds, main="Bicimad network with multidimensional scaling layout", vertex.size=8, vertex.label.cex=0.6)
```

# Decorating graph layouts

```{r, fig.width=15, fig.height=15}
V(bicimad)$size <- round(log(strength(bicimad)))
travel_time_scaled <- scale(E(bicimad)$travel_time)
E(bicimad)$width <- unlist(split(travel_time_scaled, rep(1:ncol(travel_time_scaled), each = nrow(travel_time_scaled))))+1
plot.igraph(bicimad, layout=layout_with_mds, main="Karate network with Muldimensional layout")
```

```{r, fig.width=12, fig.height=12, results='hide'}
plot.igraph(bicimad, layout = layout.norm(as.matrix(coordinates)),
            main="Bicimad network with custome layout using the latitude and the longitude",
            vertex.size=4, edge.width=0.2, edge.arrow.size=0.3, edge.arrow.width=1,
            vertex.label.cex=0.5)
```
